---
title: "hw4.2"
author: "Yue Wu"
date: "2/24/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(foreign)
library(stringr)
library(plyr)
library(reshape2)
source("pew.R")
# duplicated name 'raw' for dataset in both pew and billboard
# renaming pew data to raw1
raw1 <- raw
source("billboard.R")
```

##10.5 Exercises
#####5. What does tibble::enframe() do? When might you use it?
It converts vectors or lists to  two-column data frames.
```{r echo=TRUE, message=FALSE, warning=FALSE}
enframe(c(a = 5, b = 7))
```

##12.6.1 Exercises
Repeat the case study
```{r echo=TRUE, message=FALSE, warning=FALSE}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)

```

#####3. I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

None of the countries have multiple iso2 or iso3 codes. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
whoex3 <- select(who3, country, iso2, iso3) %>%
           group_by(country)  
n_groups(whoex3)
whoex3_2 <-  whoex3 %>% group_by(country,iso2,iso3) 
n_groups(whoex3_2)

```


#####4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.
```{r echo=TRUE, message=FALSE, warning=FALSE}
whoex4 <- who5 %>%
  group_by(country, year, sex) %>%
  dplyr::summarise(cases=sum(cases))
whoex4
  
whoex4  %>% filter(year>1990) %>%
  ggplot(aes(x = year, y = cases, group = country, color= country))   +geom_line()+ facet_wrap(~sex)+ theme(legend.position = "none")
  

```


## Tidying: Table 4 to Table 6
#### Table 4
```{r table4, ech=T}
raw1[1:10,1:7]
```
Table 4: The first ten rows of data on income and religion from the Pew Forum. Three columns, $75–100k, $100–150k and >150k, have been omitted.

#### Table 6
```{r table6, echo=T}
tb6 <- as.tibble(raw1)
tb6 %>% gather(-religion, key = "income", value = "freq") %>% arrange(religion) %>% head(n=10)
```
Table 6: The first ten rows of the tidied Pew survey dataset on income and religion. The column has been renamed to income, and value to freq.

## Tidying: Table 7 to Table 8
#### Table 7
```{r table7, echo=T}
raw[c(1:3, 6:10),1:8]
```
Table 7: The first eight Billboard top hits for 2000. Other columns not shown are $wk4, wk5,..., wk75$.


#### Table 8
```{r table8, echo=T}
tb7 <- as.tibble(raw)
tb8 <- tb7 %>% gather(key= "week", value= "rank", -year, -artist, -track, -time, -date.entered ) %>%
  arrange(artist) %>%
  select(year,artist, time, track, date=date.entered,week,rank) %>%
  filter(!is.na(rank))
head(tb8,n=15)
```
Table 8: First fifteen rows of the tidied Billboard dataset. The date column does not appear in the original table, but can be computed from $date.entered$ and $week$.